\artus -- Framework for Event-based Data Analysis in High Energy Physics

### 2015-09-09 Core/doc/latex/paper.tex
intro

Between the data taking at the detector or the generation of Monte Carlo events with the subsequent detector simulation and the showcasing of the physical results, the following analysis steps have to be performed.

The event reconstruction of the raw detector data is performed on large computing centres in the WLCG. The reconstructed data is stored in different formats at multiple sites in the WLCG. Usually, the AOD or the miniAOD format is used for physics analyses. These formats contain as much information about the events as needed to serve the general needs of the analysis groups.

This step is followed by the skimming of data sets that are relevant for a specific analysis. Both the number of events as well as the event content is reduced in order to speed up the following processing of the data. By requiring data sets for certain high level triggers and applying cuts on event quantities, the number of events is reduced. Only the event quantities, that are relevant for a given analysis, are written out. The skimming step is usually performed on grid sites where the data is located. Sometimes, the data is also transferred to sites, where free computing resources are available. The output is an analysis-dependent n-tuple format that is stored at local file servers. The size of files needed for a certain analysis ranges from $\mathcal O(10)$~GB to few~TB.

@@

The skimmed n-tuples are finally analysed by user code on local computing resources. Events are selected and categorised, high-level calculations based on the physics objects are performed and the output is either again written to smaller n-tuples or histograms. The events are usually processed with different configurations depending on the analysis needs. The output is then used for post-processing steps including the plotting of results. The post-processing does not necessarily require an event-based processing and therefore also needs on average the smallest computing resources.

>>>>

Between the data taking at the detector or the generation of simulated events with the subsequent detector simulation and the showcasing of the physical results, the following analysis steps have shown to be the best compromise in terms of turnaround time and efficiency as explained in Figure \ref{figure_analysis_steps}.

The event reconstruction of the raw detector data is performed on large computing centres in the WLCG\footnote{\url{http://wlcg-public.web.cern.ch}}, the Worldwide LHC Computing grid. The reconstructed data is stored in different formats at multiple sites in the WLCG. Usually, already a condensed format is used for physics analyses, where high-level physics objects are already reconstructed. These formats contain as much information about the events as needed to serve the general needs of the analysis groups.

This step is followed by the selection of data sets that are relevant for a specific analysis. This step is usually being referred to as skimming. Skimming means, that only those events that are interesting for the analysis are selected, e.g. by requiring only data sets for certain high level triggers or the presence of certain physical objects with desired kinematical properties. Only the event quantities, that are relevant for a given analysis, are written out. Since the WLCG follows the concept of data locality, the skimming step is usually performed on grid sites where the data is located. With rising network bandwith over the last few years, data locality is weakened up and techniques arose that also allow the access of data on remote sites. The output of the skimming is an mostly analysis-dependent n-tuple format that is small enough to be at smaller file servers, often owned and operated by a single intitute. The size of files needed for a certain analysis ranges from $\mathcal O(10)$~GB to few~TB.

@@

The skimmed n-tuples are finally processed by user code on resources indipendent from the WLCG. The analysis performs further events selectetion and categorization. High-level calculations based on the physics objects are performed and the output is either again written to smaller n-tuples or histograms. The events are usually processed with different configurations depending on the analysis needs. The output is then used for post-processing steps including the plotting of results. The post-processing does not necessarily require an event-based processing and therefore also needs on average the smallest computing resources.
